---
title: "Modeling Cars -- Bayes Style"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(ggformula)
library(mosaic)
library(readr)      
library(brms)
library(bayesplot)

load("Cars_brm.Rda")

knitr::opts_chunk$set(echo = FALSE)
options(scipen = 100)
options(digits = 3)
theme_set(theme_bw())
```


## Introduction



### Less Volume, More Creativity

The point here is not to give a comprehensive introcution to either 
the brms package or to Bayesian modeling, but to show that because the brms package
uses a familiar interface, one can move from least squares/maximum likelihood fits
of linear models to Bayesian linear models with relative ease.

### The Data

In this tutorial we will demonstrate how to fit a multivariate bayesian
regression model to describe the relationship between the retail price of used
GM cars.  
We include several variables in this dataset that might influence the
`Price` of a car:

* Mileage (number of miles the car has been driven)
* Make (Buick, Cadillac, Chevrolet, Pontiac, SAAB, Saturn)
* Type (Sedan, Coupe, Hatchback, Convertible, or Wagon)
* Cyl (number of cylinders: 4, 6, or 8)
* Liter (a measure of engine size)
* Doors (number of doors: 2, 4)
* Cruise (1 = cruise control, 0 = no cruise control)
* Sound (1 = upgraded speakers, 0 = standard speakers)
* Leather (1 = leather seats, 0 = not leather seats)

We start by reading in the data.

```{r intro, message = FALSE, echo = TRUE}
library(readr)
GMCars1 <- read_csv("data/GMCars1.csv")   # Reads the data 

head(GMCars1, 5)    # Shows the first 5 rows of the data 
```

## Bayesian linear models with brms

The `brms` package allows us to fit Bayesian linear models relatively easily.  To use it,
you need to do some initial configuration and install Stan, which does the heavy lifting.
Stan requires that you have C/C++ compiler on your machine.

### Fitting with brms

We can use the same model formula as with `lm()` or `glm()` or (`lmer()` or `glmer()`). 
This makes it easy to move to Bayesian models.

The `brm()` function converts this information
to Stan code, and Stan uses a Hamilton Monte Carlo algorithm to attempt to draw samples
from the posterior distribution.  The whole process can take some time, but isn't that 
complicated.

  * Compiling the model typically take 30-60 seconds.
  * Generating posterior samples can take seconds to hours to days depending on the complexity
  of the model and of the data.  (For linear models with only a few parameters, it typically
  doesn't take very long.  For complicated heierarchical models or models with many parameters,
  it can take longer.)
  
To save some time, the following models have been fit already:


```{r, eval = FALSE}
library(brms)
Cars.brm1 <- brm(Price ~ Mileage,        data = GMCars1)
Cars.brm2 <- brm(Price ~ Mileage + Make, data = GMCars1) 
Cars.brm3 <- brm(Price ~ Mileage * Make, data = GMCars1) 
```

### Posterior

Let's take a look at the posterior distribution for each parameter. 
The `plot()` function automates this for us.
(The plots on the right are diagnostics to show how well the HMC algorithm is converging.)

```{r, plot, exercise = TRUE}
plot(Cars.brm2)
```

`stanplot()` provides a graphical representation of the parameter estimates:

```{r, stanplot, exercise = TRUE}
stanplot(Cars.brm2)
```

Alternatively, we can get the posterior samples in a tidy data frame and do
whatever we like with them.

```{r, poster-df, exercise = TRUE}
posterior_samples(Cars.brm2)
gf_density(~ b_Mileage, data = posterior_samples(Cars.brm2))
posterior_interval(Cars.brm1, prob = 0.95)   # HDPIs for each parameter
posterior_interval(Cars.brm1, prob = 0.95, pars = "b_Mileage")  
gf_point(b_Mileage ~ b_MakePontiac, data = posterior_samples(Cars.brm2))
```

There are many other functions that can be used to obtain information about our model:

```{r methods, echo = TRUE}
methods(class = "brmsfit")
```

## Priors

So what priors did we just use?  Let's ask.

```{r, prior-summary, exercise = TRUE}
prior_summary(Cars.brm1)  # try the other models, too
```

Where no prior is listed, `brms` is using a flat, improper prior.  (Basically a "uniform 
distribution over the real line.)  This is rarely the best prior to use.  Typically we
want to use something that is unimodal (often centered at 0 or some other "natural value", 
often with heavy tails to keep from influencing the model too much).  But to choose a good
prior one must know something about the data and the model.


Here are some examples with a custom prior.  These are not intended to be especially 
good priors, just to demonstrate how a prior is set.  Many other distributions besides
a normal distribution can be used.


```{r, eval = FALSE}
library(brms)
Cars.brm1a <- 
  brm(Price ~ Mileage,        data = GMCars1,
      prior = set_prior("normal(0, 2)")
  )
Cars.brm2a <- 
  brm(Price ~ Mileage + Make, data = GMCars1, 
      prior = set_prior("normal(0, 2)")
  )  
Cars.brm2b <- 
  brm(Price ~ Mileage + Make, data = GMCars1, 
      prior = c(
        set_prior("normal(0, 2)",  coef = "Mileage"),
        set_prior("normal(0, 50)", coef = "MakePontiac"),
        set_prior("normal(0, 50)", coef = "MakeSAAB")
      ))  
Cars.brm3a <- 
  brm(Price ~ Mileage * Make, data = GMCars1,
      prior = c(
        set_prior("normal(0, 2)",  coef = "Mileage"),
        set_prior("normal(0, 50)", coef = "MakePontiac"),
        set_prior("normal(0, 50)", coef = "MakeSAAB"),
        set_prior("normal(0, 2)",  coef = "Mileage:MakePontiac"),
        set_prior("normal(0, 2)",  coef = "Mileage:MakeSAAB")
      ))  
```

Feel free to explore this models here

```{r, explore, exercise = TRUE}
fixef(Cars.brm2)
fixef(Cars.brm2a)
fixef(Cars.brm2b)
prior_summary(Cars.brm2a)
```


## What does Stan look like?

brms only handles a general class of (generalized, heierarchical) linear models.  That covers
a lot of useful cases.  But Stan is even more general.  If you are curious to see what native
Stan code looks like, you can extract the Stan code from a model fit by brms.
This can be useful

* to make sure the model is doing what you intend,
* as a stepping stone toward learning Stan


```{r stan-code, exercise = TRUE}
# try other models if you like
stancode(Cars.brm2)
```


## Where do I learn more

Here is some online documentation for

* [brms](https://github.com/paul-buerkner/brms)
* [Stan](https://mc-stan.org/)
* [rethinking](https://github.com/rmcelreath/rethinking) -- a package that accompanies
*Statistical Rethinking* (McElreath) and provides a simplified but still quite
flexible interface to Stan.

JAGS is another system for fitting Bayesain models, and there are R packages that interface
with that as well.  

* [Using JAGS via R](http://www.jkarreth.net/files/bayes-cph_Tutorial-JAGS.pdf)

For more info about Bayesian things in R, see that [CRAN task view](https://cran.r-project.org/web/views/Bayesian.html)

