---
title: "Introduction to RMarkdown Files_Part2"
author: "Shonda Kuiper"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Introduction

This is Part 2 of an introduction to R Markdown. In this document, you will be asked to read and analyze several datasets using data from the [Statistically Grounded Game](http://web.grinnell.edu/individuals/kuipers/stat2labs/CoffeeTruck.html). In this game, students collect sales data on a small coffee business. Players can design experiments to determine how to maximize profit within the game by adjusting various factors, such as location, time of day, price, and type of music. The goal is to design a business plan that will earn the most sales and make the largest amount of profits.

We start by using the `readr` package to read our data, then `mosaic` and `ggformula` packages to view the data and conduct some initial statistical analysis. In later sections, we use two additional packages to make modifications to the data. We start by requiring all packages that we will be using in this R Markdown document.

```{r, message = FALSE, warning = FALSE}
library(stringr)    # A package that allows us to manipulate text  
library(dplyr)      # A package that is used for data wrangling
library(ggformula)  # A package to create graphs
library(mosaic)     # A package for introductory statistics calculations
library(readr)      # A package to read data

Coffee <- read_csv("data/Coffee.csv")   # Reads the data 
head(Coffee, 5)                         # Shows the first 5 rows of the data 
```

This is a cleaned and organized subset of a full dataset that you will view later. The data includes the following variables:

* Day: the day (1 - 40) when the data was collected. Since this is a subset of the full experiment, not every day within this data.
* Location: where the coffee was sold (City Park or in the Business District)
* Price: The cost of each cup of coffee (every cup was sold for $3 in this sample)
* Time: The time of day when the coffee was sold (Morning or Afternoon)
* Temperature: The average daily temperature during the Day/Time coffee was sold.
* Sales: The number of cups sold during that Day/Time.
* Income: The income earned during that time (Sales * Price)

## 2. Initial Investigation: The Relationship Between Location and Sales.

We will start with a few commands from the `mosaic` package to view our data. Recall that `mosaic` uses a common structure for all its commands:

<center>
<h3><strong><span class="invboxed">goal</span> ( <span class="invboxed">y</span> ~ <span class="invboxed">x</span> , data = <span class="invboxed">mydata</span> )</strong> 
</h3></center>



```{r fig.height = 3, fig.width = 6}
favstats(Sales ~ Location, data = Coffee)

tally(~ Location, data = Coffee)

gf_boxplot(Sales ~ Location, data = Coffee) %>%
gf_labs(title = "Figure 1", caption = "Boxplot of Coffee Truck Sales by Location")
```


Without carefully evaluating the data, some people may decide to conduct a t-test to compare these two groups. Our null and alternative hypotheses are: 

<center> $H_0: \mu_B = \mu_P$  vs.  $H_1: \mu_B \neq \mu_P$ </center>

```{r}
t.test(Sales ~ Location, data = Coffee)
```

We see that the t.test indicates that there is no evidence of differences in `Sales` based upon `Location`. However, let's view the data a little more carefully. In particular, look at the influence of a third variable, `Time`.

```{r fig.height = 3, fig.width = 6}
gf_boxplot(Sales ~ Location, data = Coffee, color = ~Time) %>%
gf_labs(title = "Figure 2", caption = "Boxplot of Coffee Truck Sales by Location and Time of Day")

```

The plot clearly indicates that there is an **Interaction Term** that should be included in our analysis, since the effect of `Location` on `Sales` is influenced by a third variable, `Time`.


```{r fig.height = 3, fig.width = 6}
summary(aov(Sales ~ Location * Time, data = Coffee))
```

## 3. Cleaning Data

If you have not worked with R packages to clean and organize data, the following commands may be somewhat confusing at first. In particular, we are using the `stringr` and `dplyr` packages in this section. *It is not essential that you understand what each function is doing at this time.* The goal of this section is simply to share how the source data was cleaned and modified to prepare the `Coffee` data set used earlier.

```{r message = FALSE}
# Read in the full dataset:
CoffeeFull <- read_csv("data/CoffeeTruck.csv")
dim(CoffeeFull)
tally(`Time of Day` ~ Price, data = CoffeeFull)
```

Notice that the `Time of Day` variable includes four different times of day. Also the `Price` variable has four levels and are represented as characters with a `$` in front of each number.

```{r message = FALSE}
# The `str_sub` command extracts the character in the second position for the Price variable:
CoffeeFull <- mutate(CoffeeFull, Price = as.numeric(str_sub(Price, 2)))
dim(CoffeeFull)

# Restrict data to only the samples of interest
Coffee1 <- filter(CoffeeFull, `Time of Day` =="Afternoon" | `Time of Day` =="Morning")
Coffee1 <- filter(Coffee1, Price == 3)
#dim(Coffee)

# Select specific columns of data, and abbreviate some column titles
Coffee1 <- select(Coffee1, Day, Location, Price, Time = `Time of Day`, Temperature, Sales, Income = `Gross Income`)
#dim(Coffee)

```
Now that we have the `Coffee` data (Now called `Coffee1` in the format we like can run the same analysis as in the previous section to get identical results.

```{r fig.height = 3, fig.width = 6}
gf_boxplot(Sales ~ Location, data = Coffee1, color = ~Time) %>%
gf_labs(title = "Figure 3", caption = "Boxplot of Coffee Truck Sales by Location and Time of Day")

summary(aov(Sales ~ Location * Time, data = Coffee1))
```



## Creating Reproducible Results

#### Reproducing Data 1

Now we will determine if these results are reproducible. Since this data was collected for a class project, many students used the same methods to collect their own unique simple random sample. A second dataset, called `CoffeeTruck2.csv` provides a second dataset. 

To evaluate this data, and determine if we get similar results for both samples, simply replace 

line 90: `CoffeeFull <- read_csv("data/CoffeeTruck.csv")`  with

line 90: `CoffeeFull <- read_csv("data/CoffeeTruck2.csv")` and Knit the document.

* Are the boxplots and ANOVA for Sections 2 and 3 similar? 
* Did you come to the same conclusions with both datasets? 
* Do you think sample size had any influence on the reproducibility of this study?

R Markdown allows us to quickly identify whether any any data points were removed or modified. In addition, R Markdown makes it extremely easy to repeat an earlier analysis. After the data is collected, only one line of code needs a small change, `CoffeeFull <- read_csv("data/CoffeeTruck.csv")` and the analysis is completely redone.

While it initially may take a little longer to analyze data with R Markdown files, there is great benefit in being able to quickly validate your data cleaning and statistical models, as well as creating reproducing results. 




#### Optional: Reproducing Data 2

If you want to conduct your own validation of reproducibility, you can collect your own data, clean the data in a similar manner, and determine whether you get similar results. Here are the steps you need to complete:

1. Go to the [Statistically Grounded Game](http://statgames.tietronix.com/statisticallygrounded/statisticallygrounded_webgl/index.html).
     * Check to save your data
     * Enter a name (you do not need your actual name, just any PlayerID that you will remember)
     * Enter the Group Name: JMM2019
2. Read through the brief instructions and select:
     * Location: Business District and City Park
     * Music: No Music
     * Price: 2 per cup, 3 per cup, 4 per cup, and 5 per cup
     * Location: Morning, Lunch, Afternoon, and Evening
     * Design Type: Random Sample
     * Number of Repetitions: 5
     * Start Sim
     * Get all data
3. You now have your own unique simple random sample of data that follows the same data collection procedure as the original data. Now you can select:
     * Continue
     * Get Data
4. Download the .csv file and save it on your desktop as `CoffeeTruck3`. Now use the `Upload` button in the `Files` window of Rstudio.cloud and save your data as `CoffeeTruck3` in the `data` folder.. 
5. Change the following lines of code to `eval = TRUE`, `echo = TRUE` and change `data/CoffeeTruck.csv` to `data/CoffeeTruck3.csv`. 
6. Run the lines of code below to conduct your own reproducible analysis. Did your ANOVA results and boxplot agree with those in Section 2?


```{r message = FALSE, eval = FALSE, echo = FALSE}
# Read in the full dataset:
CoffeeFull <- read_csv("data/CoffeeTruck.csv")
dim(CoffeeFull)
tally(`Time of Day` ~ Price, data = CoffeeFull)
```


```{r message = FALSE, eval = FALSE, echo = FALSE}
# The `str_sub` command extracts the character in the second position for the Price variable:
CoffeeFull <- mutate(CoffeeFull, Price = as.numeric(str_sub(Price, 2)))
dim(CoffeeFull)

# Restrict data to only the samples of interest
Coffee1 <- filter(CoffeeFull, `Time of Day` =="Afternoon" | `Time of Day` =="Morning")
Coffee1 <- filter(Coffee1, Price == 3)
#dim(Coffee)

# Select specific columns of data, and abbreviate some column titles
Coffee1 <- select(Coffee1, Day, Location, Price, Time = `Time of Day`, Temperature, Sales, Income = `Gross Income`)
#dim(Coffee)

```


```{r fig.height = 3, fig.width = 6, eval = FALSE, echo = FALSE}
gf_boxplot(Sales ~ Location, data = Coffee1, color = ~Time) %>%
gf_labs(title = "Figure 3", caption = "Boxplot of Coffee Truck Sales by Location and Time of Day")

summary(aov(Sales ~ Location * Time, data = Coffee1))
```


