---
title: "Roadless America"
date: "JMM 2019"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(learnr)
library(mosaic)
theme_set(theme_bw())
```


## How much of the continental US is roadless?

### Definition of roadless

Let's define roadless as more than one mile from a raod.  Suppose we want to estimate
what fractions of the continental US is roadless.

### How precise an estimate do we want?

The worst case margin of error for a given sample size is easy to calculate.
By plugging in different values of $n$, we can see how large our sample size must
be for a desired level of precision.  
(Note: If we have a rough idea of the true proportion and it
is not near 50%, we can improve the estimate by choosing something we are quite certain is
closer to 0.5 than the true value.)

Experiment with different values of `n` to determine a sample size that will give us a 95%
confidence interval with margin of error at most $\pm 5$%, $\pm3$%, or $\pm1$%. 
(Are we willing to put in enough effort to estimate to $\pm1$%?)

```{r sample-size, exercise=TRUE}
n <- 100
ME <- 2 * sqrt(0.5 * 0.5 / n); ME
```

### Graphcial Appraoch

Or we can explore graphically.  Adjust `xlim` and `yintercept` to determine the sample size
required for a margin of error of $\pm$ 1%.

```{r graphcial-sample-size, exercise = TRUE, fig.width = 8, fig.height = 3}
library(mosaic)
gf_fun(2 * sqrt(0.5 * 0.5 / n) ~ n, xlim = c(100, 2000)) %>%
  gf_hline(yintercept = 0.03, color = "red")
```

## Sampling 

We can use `rgeo()` to sample random points on the globe.  By adding limits to longitude and 
latitude, we can restrict ourselves to a region just a bit larger than the continental US.

```{r locations, exercise = TRUE}
Locations <- rgeo(4, latlim = c(25,50), lonlim = c(-65, -125))
Locations
```

### Classifying locations

The sampled points will be in three categories:

* Within 1 mile of a road
* Not within 1 mile of a road
* Not in contentental US

It would be fun to visit all these places, but it is cheaper to use a map.

The map below is interactive.  Zoom in on each point to see if it is roadless.

```{r leaflet, exercise = TRUE}
Locations <- rgeo(10, latlim = c(25,50), lonlim = c(-65, -125))
leaflet_map(Locations, radius = 1, unit = "miles")
Locations
```

## Collecting and Analysing Data

A google spreadsheet or google form (or other online data collection services) can be used
to assemble data from many students in parallel.  `binom.test()` can be used to compute
a confidenc interval.

```{r binom-test, exercise = TRUE}
# adjust the numbers 100 and 300 to reflect actual sample 
binom.test(100, 300) %>% 
  confint()
```